Adversarial Training of Neural Networks\\ against Systematic Uncertainty

### 2016-08-30 tex/paper.tex
added note
save
some tex cleanup

% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

\usepackage[final]{nips_2016}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{color}

\usepackage{algorithm}
\usepackage{algpseudocode}

@@

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Gilles Louppe \\
  New York University\\
  \texttt{g.louppe@nyu.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

\begin{document}
% \nipsfinalcopy is no longer used

@@

@@

\begin{proposition}\label{prop:1}

Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = \frac{1}{|{\cal Z}|}$
for all $z_l$, then $f$ is a pivotal quantity.

\end{proposition}

\begin{proof}

Let us first recall that the cross-entropy for distributions $p$ and $q$ is
minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
the approximate conditional distribution of the nuisance
$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\smash{\hat\theta}_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$ are therefore such that $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$.

@@

\begin{proposition}\label{prop:2}

If there exists a saddle point $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$
for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
$E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - \log |{\cal Z}|$, then
$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier  and a pivotal quantity.

\end{proposition}

\begin{proof}

For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$, in which case $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$ and $L_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
\log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the log-loss $L_f$
    reduces to its minimum value $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z_l|f(X;\smash{\hat\theta}_f)) = \frac{1}{|{\cal Z}|}$ for all $z_l \in {\cal
   Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
   -\sum_{z_l \in {\cal Z}} p(z_l|f(X;\smash{\hat\theta}_f)) \log  p(z_l|f(X;\smash{\hat\theta}_f)) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
\end{itemize}
Accordingly, the second condition implies that $r(f(X;\smash{\hat\theta}_f) ;
\smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
\end{proof}

\glnote{We should further discuss that in practice, the equality may never hold.
We should discuss in which circumstances. In such case, the pivotal quantity
constraint can however be enforced by outweighting the $L_r$ term, resulting in a
trade-off between classifier optimality and pivotality.}

@@

@@

@@

\bibliographystyle{ieeetr}
{\small
\bibliography{bibliography.bib}}

>>>>

\PassOptionsToPackage{numbers, compress}{natbib}

\usepackage[final]{nips_2016}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Gilles Louppe \\
  New York University\\
  \texttt{g.louppe@nyu.edu} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  %% \texttt{email} \\
}

@@

\begin{document}

@@

% ==============================================================================

@@

\begin{proposition}\label{prop:1}
Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = \frac{1}{|{\cal Z}|}$
for all $z_l$, then $f$ is a pivotal quantity.
\end{proposition}

\begin{proof}
Let us first recall that the cross-entropy for distributions $p$ and $q$ is
minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
the approximate conditional distribution of the nuisance
$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\smash{\hat\theta}_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$ are therefore such that $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$.

@@

\begin{proposition}\label{prop:2}
If there exists a saddle point $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$
for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
$E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - \log |{\cal Z}|$, then
$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier  and a pivotal quantity.
\end{proposition}

\begin{proof}
For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$, in which case $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$ and $L_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
\log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the log-loss $L_f$
    reduces to its minimum value $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z_l|f(X;\smash{\hat\theta}_f)) = \frac{1}{|{\cal Z}|}$ for all $z_l \in {\cal
   Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
   -\sum_{z_l \in {\cal Z}} p(z_l|f(X;\smash{\hat\theta}_f)) \log  p(z_l|f(X;\smash{\hat\theta}_f)) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
\end{itemize}
Accordingly, the second condition implies that $r(f(X;\smash{\hat\theta}_f) ;
\smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
\end{proof}

@@

\glnote{We should further discuss that in practice, the equality in Prop 2 may never hold.
We should discuss in which circumstances. In such case, the pivotal quantity
constraint can however be enforced by outweighting the $L_r$ term, resulting in a
trade-off between classifier optimality and pivotality.}

@@

\glnote{Other applications: removing implicit bias in data (e.g. gender bias).}

% ==============================================================================

@@

\bibliographystyle{ieeetr}
\bibliography{bibliography.bib}
